---
title: "Soma vs Nuclei Size"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(feather)

RotateCoords <- function(xy, rot.angle = 0) {
  rotm <- matrix(c(
      cos(rot.angle),
      sin(rot.angle),
      -sin(rot.angle),
      cos(rot.angle)), ncol = 2)
  xy.rot <- t(rotm %*% (t(xy) - c(xy[1, 1], xy[1, 2])) + c(xy[1, 1], xy[1, 2]))
}
```

## Size of nuclei and soma 

We aim to measure the size of soma and nuclei for cells in different layers of the mouse visual cortex. Cells labeled by layer-enriched Cre lines crossed to Ai14 are imaged on confocal microscope. Maximum intensity projection of 5-um optical z-stacks (at 1-um intervals) from confocal images are analyzed on CellProfiler to segment soma and their associated nuclei.

## Plotting soma area vs nuclei area for Cre-lines.
```{r plot-nuc-prop-vs-cre, echo=FALSE, fig.width=4, fig.height = 4}
# Load data
area.dat <- read.csv(file = "../data/nuc_soma_area_cre_lines.csv")

area.dat$n_vol_est <- area.dat$n_area^(3/2)
area.dat$s_vol_est <- area.dat$s_area^(3/2)
area.dat$nuc_soma_ratio_vol <- area.dat$n_vol_est / area.dat$s_vol_est

for (cre1 in unique(area.dat$cre)) {
  area.subset <- subset(area.dat, cre == cre1)
  lm1 <- lm(I(n_vol_est) ~ I(s_vol_est) + 0, data = area.subset)
  print(paste(cre1, round(coef(lm1), 2)))
  # print(summary(area.subset$nuc_soma_ratio_vol))
}

aov1 <- aov(nuc_soma_ratio_vol ~ cre, data = area.dat)
summary(aov1)
print(TukeyHSD(aov1))

g.scatter <- ggplot(area.dat, aes(x = s_area, y = n_area, color = cre)) +
  # facet_wrap(~ cre, ncol = 1) +
  geom_point(size = 0.5) +
  geom_smooth(method = "lm", formula = "y ~ 0 + x", se = FALSE, size = 0.3) + 
  xlab(expression(paste("Soma area (", mu*m^2, ")"))) +
  ylab(expression(paste("Nucleus area (", mu*m^2, ")"))) +
  theme_bw()
plot(g.scatter)
ggsave(g.scatter, file = "../output/nuc_prop_cre_scatter.pdf", 
       height = 2.5, width = 4)


# Cre-line nuclear proportions estimated based on cell type mixture
nuc.prop.expected <- data.frame(cre = c("rbp4", "nr5a1", "scnn1a_tg3"),
                                expr_nuc_prop = c(0.29, 0.49, 0.50),
                                intron_nuc_prop = c(0.46, 0.72, 0.74))
g.vio <- ggplot(area.dat, aes(x = cre, y = nuc_soma_ratio_vol)) +
  geom_violin(size = 0.3) +
  geom_jitter(alpha = 0.2, width = 0.3, size = 0.5, show.legend = FALSE) +
  geom_segment(data = nuc.prop.expected, color = "blue", size = 0.8,   
               mapping = aes(x = cre, xend = cre, y = expr_nuc_prop, 
                             yend = intron_nuc_prop)) +
  stat_summary(fun.y = mean, geom = "point", size = 2.5, show.legend = FALSE) +
  xlab("Cre-line") +
  ylab("Nuclear proportion") +
  theme_bw()
plot(g.vio)
ggsave(g.vio, file = "../output/nuc_prop_cre_violin.pdf", 
       height = 2.5, width = 2.5)

```

```{r nuc_prop_vs_depth, fig.width=6, fig.height = 4}
# Load dat
nuc.soma <- read.csv(file = "../data/nuc_soma_area_wt.csv")


nuc.ratio.df <- data.frame()
for (i in seq(1, nrow(nuc.soma), 2)) {
  x.mean <- mean(nuc.soma$CenterX[c(i, i + 1)])
  y.mean <- mean(nuc.soma$CenterY[c(i, i + 1)])
  # Check that nuc/soma outlines are paired (centers aligned)
  x.cv <- sd(nuc.soma$CenterX[c(i, i + 1)]) / mean(nuc.soma$CenterX[c(i, i + 1)])
  nuc.area <- nuc.soma$Area[i + 1]
  soma.area <- nuc.soma$Area[i]
  nuc.area.ratio <- nuc.area / soma.area
  nuc.vol.ratio <- nuc.area.ratio^(3/2)
  nuc.ratio.df <- rbind(nuc.ratio.df, cbind(x.mean, y.mean, x.cv, nuc.area, soma.area,
                                            nuc.area.ratio, nuc.vol.ratio))
}

# Rotate coords
alpha = -68  # Measured rotation
xy <- cbind(nuc.ratio.df$x.mean, -nuc.ratio.df$y.mean)
xy.rot <- RotateCoords(xy, rot.angle = alpha)

nuc.ratio.df <- cbind(nuc.ratio.df, xy.rot)
nuc.ratio.df$depth_from_pia <- max(nuc.ratio.df$`2`) - nuc.ratio.df$`2` + 30


g.prop.v.depth <- ggplot(nuc.ratio.df, aes(y = -depth_from_pia, x = nuc.ratio.df$nuc.vol.ratio,
                               size = soma.area^(3/2), color = soma.area^(3/2))) +
  geom_point(show.legend = FALSE) +
  xlab("Nuclear proportion") +
  ylab(expression(paste("Depth from pia (", mu*m, ")"))) +
  # ggtitle("Nuclear proportion of soma") +
  scale_color_gradient(low = "grey80", high = "black") +
  theme_bw(base_size = 18)
plot(g.prop.v.depth)
ggsave(g.prop.v.depth, file = "../output/nuc_prop_vs_depth.pdf", 
       height = 4, width = 4)

```

